\documentclass[twoside]{article}
\usepackage{rawfonts}
\IfFileExists{times.sty}{\usepackage{times}}{\@missingfileerror{times}{sty}}
\oddsidemargin 0.1in
\evensidemargin -0.1in
\def\textwidth{6.375 in}

\title{Installing the WFDB Software Package}
\author{George B. Moody\\
Harvard-MIT Division of Health Sciences and Technology, Cambridge, MA, USA}
\date{}

\begin{document}
\setcounter{page}{121}

\maketitle

\section*{Summary}
This appendix briefly describes how to install the WFDB Software Package on a
new system.  The package includes C-language sources for the WFDB library and
for the applications described in this manual; \LaTeX{}, {\tt texinfo}, and
{\tt troff} sources for this and other manuals; and a one-minute sample record
({\tt 100s}).

\section*{Obtaining the WFDB Software Package}
The latest version of the WFDB Software Package can always be downloaded in
source form from PhysioNet and its mirrors (see {\tt
http://www.physio\-net.org/physio\-tools/wfdb.shtml}); binaries for popular
operating systems and development snapshots, as well as quick-start guides
including installation notes for popular operating systems, are also usually
available there.

\subsection*{UNIX, GNU/Linux, and similar operating systems}

Before beginning the installation of the WFDB Software Package, obtain and
install the {\tt libwww} package from {\tt http://www.w3.org/Library/} or from
{\tt http://\-www.physio\-net.org/physio\-tools/libwww/}.  This package is
provided with most current versions of GNU/Linux.  (If you have a program
called {\tt libwww-config}, then {\tt libwww} is installed already.)  You may
omit this step if you do not wish to have NETFILES support.

Also download and install the {\tt XView} software packages from {\tt
http://www.\-physio\-net.org/\-physio\-tools/xview/} if you wish to use WAVE.
Sources are available, as are binaries for several versions of GNU/Linux.  If
you are using SunOS or Solaris, XView binaries are available in the Open Look
Software Development package and may be installed already.  (If you have a
program called {\tt textedit}, then {\tt XView} is installed already.) If you
are able to use an existing set of binaries, these are recommended, since the
sources may take a {\em long} time to compile.  Be sure that the directory
containing {\tt textedit}, usually {\tt /usr/openwin/bin}, is in your {\tt
PATH}.  You may omit this step if you do not wish to use {\em WAVE}.

If you have downloaded the software from PhysioNet or another source, you
will have a {\tt gzip}-compressed {\tt tar} archive.  Unpack it using
the commands:

\begin{verbatim}
gzip -d wfdb.tar.gz
tar xfv wfdb.tar
\end{verbatim}

(If you have GNU {\tt tar}, as on GNU/Linux, you can combine these into a
single command: {\tt tar xfvz wfdb.tar.gz}.)

This will create a directory with a name of the form {\tt wfdb-}{\em m.n.r},
where {\em m.n.r} is the version number of the included WFDB library (e.g.,
{\tt 10.2.1}).  Enter this directory.

You should now be ready to configure, compile, and install the software, using
the commands:

\begin{verbatim}
./configure
make install
\end{verbatim}

The {\tt ./configure} command asks where you wish to install the package.
If you accept the default ({\tt /usr}), you will need root permissions when
runnning {\tt make install}.  If you choose another location, follow the
instructions given by {\tt configure} for setting your {\tt PATH} and
{\tt LD\_LIBRARY\_PATH} environment variables.

Depending on the speed of your system and of your C compiler, {\tt make} will
generally require between 1 and 10 minutes.

\subsection*{MS-Windows}

If you have not already done so, install the Cygwin development environment
(freely available from {\tt http://\-sources.\-red\-hat.\-com/\-cygwin/}).  This includes
{\tt gcc} (the GNU C/C++ compiler) as well as a comprehensive assortment of
other Unix utilities ported to MS-Windows.  Accept the defaults suggested by
the installer.

\emph{Important:} Although you may be able to compile the WFDB software
package using a proprietary compiler, this is \emph{not supported}.  The
{\tt Makefile.dos} files in several of the subdirectories of the package's
source tree can be used with the {\tt make} utilities provided with most
commercial C compilers, although you will need to customize them for your
compiler. Your feedback is appreciated.

Before beginning the installation of the WFDB Software Package, obtain and
install the {\tt libwww} package from {\tt http://www.w3.org/Library/} or
from {\tt http://\-www.physio\-net.org/physio\-tools/libwww/}.  You may
omit this step if you do not wish to have NETFILES support.

Download
{\tt http://www.physio\-net.org/physio\-tools/binaries/win\-dows/bin/which.exe}
and put it into a directory in your PATH.  (This utility is needed by
{\tt configure} in a later step.  The sources for {\tt which.exe} are available
within {\tt http://www.physio\-net.org/physio\-tools/util\-ities/}.)

Open a Cygwin terminal window (the Cygwin installer will have added this to
your MS-Windows start menu).  Perform the remaining steps by typing the
commands given below into the terminal window.

Check that {\tt which} and {\tt gcc} are accessible by typing the command:

\begin{verbatim}
which gcc
\end{verbatim}

The output of this command should be:

\begin{verbatim}
/usr/bin/gcc
\end{verbatim}

If you don't see this output, repeat the steps above as necessary to correct
the problem before continuing.

Unpack the {\tt gzip}-compressed {\tt tar} archive you downloaded earlier,
using the {\tt tar} command included with the Cygwin package:

\begin{verbatim}
tar xfvz wfdb.tar.gz
\end{verbatim}

If your browser decompressed the file during the download, use this command
instead:

\begin{verbatim}
tar xfv wfdb.tar
\end{verbatim}

This will create a directory with a name of the form {\tt wfdb-}{\em m.n.r},
where {\em m.n.r} is the version number of the included WFDB library (e.g.,
{\tt 10.2.0}).  Enter this directory.

You should now be ready to configure, compile, and install the software, using
the commands:

\begin{verbatim}
./configure
make install
\end{verbatim}

If you have Microsoft or Turbo C or C++, and a Microstar Laboratories DAP 1200-
or 2400-series analog interface board, you can compile {\tt sample} (a program
for creating database records from analog signals, and for replaying them in
analog form).  To do so successfully, you must first have installed the
Microstar {\tt \#include} files and DAP interface library on your system.
Specifically, files {\tt c\_lib.c}, {\tt clock.h}, and {\tt ioutil.h} must be
installed in your {\tt include} directory, and the version of the file
{\tt cdapl.lib} that is compatible with your compiler must be installed in a
directory in which libraries are found by your linker.  Read and customize
{\tt lib/Makefile.dos} and {\tt app/Makefile.dos} as appropriate for your
compiler, and use your compiler's {\tt make} utility to generate {\tt wfdb.lib}
and then {\tt sample.exe}.

\subsection*{Other systems}

Copy the contents of {\tt http://www.\-physio\-net.\-org/physio\-tools/wfdb/}
to your hard disk.  Note that the text files are in UNIX format (i.e., lines
are terminated by ASCII line-feed characters only).  If your system expects
text files in MS-DOS format (with both a carriage return and a line-feed at the
end of each line; VMS is one such system), use {\tt u2d.exe} (available from
PhysioNet) to reformat the text files under MS-DOS.  If your system is a
Macintosh (which expects that lines are terminated by carriage returns only),
you will have to reformat the text files yourself, which may be done under
MS-DOS on a PC using {\tt u2m.exe}, or on a Macintosh using third-party
software.

The WFDB Software Package is written in highly portable C, and (with the
exception of a few MS-DOS or UNIX-specific display or data-acquisition
programs) should be easy to compile with any K\&R or ANSI C compiler.
The UNIX and MS-DOS {\tt make} description files ({\tt Makefile}
and {\tt Makefile.dos} in {\tt wfdb} and in each of its subdirectories)
should get you started.
\end{document}
